---
title: "p8105_hw3_tal2145"
author: Troy Layouni
date: 2019-10-08
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Loading packages**

```{r}
library(tidyverse)
library(ggridges)
```

## Problem 1

**Problem 1, Part 1: Loading Instacart data**

```{r}
library(p8105.datasets)
data("instacart")
```


**Problem 1, Part 2: Number of aisles and the aisle with the most items ordered**

```{r}
instacart_aisles = 
  instacart %>% 
  group_by(aisle) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count)) 

instacart_aisles
```

There are 134 aisles in the instacart dataset and the aisle that the most items are ordered from is the fresh vegetables. 

**Problem 1, Part 3: Plotting the number of items ordered in each aisle, limited to aisles with more than 100000 items ordered**

```{r}
instacart_aisles %>% 
  filter(count > 10000) %>% 
  
ggplot(aes(x = aisle, y = count)) +
  geom_col() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(
   title = "Number of Items order by Aisle",
   x = "Aisle",
   y = "Number of Items Ordered",
   caption = "Data from Instacart"
  ) +
  viridis::scale_color_viridis(
    name = "aisle", 
    discrete = TRUE
)

```
    




