---
title: "p8105_hw3_tal2145"
author: Troy Layouni
date: 2019-10-08
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Loading packages**

```{r}
library(tidyverse)
library(ggridges)
```

## Problem 1

**Problem 1, Part 1: Loading Instacart data**

```{r}
library(p8105.datasets)
data("instacart")
```


**Problem 1, Part 2: Number of aisles and the aisle with the most items ordered**

```{r}
instacart_aisles = 
  instacart %>% 
  group_by(aisle) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n)) 

instacart_aisles
```

There are 134 aisles in the instacart dataset and the aisle that the most items are ordered from is the fresh vegetables. 

**Problem 1, Part 3: Plotting the number of items ordered in each aisle, limited to aisles with more than 10,000 items ordered**

```{r}
instacart_aisles %>% 
  filter(n > 10000) %>% 
  
ggplot(aes(x = aisle, y = n)) +
  geom_col() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(
   title = "Number of Items Ordered by Aisle",
   x = "Aisle",
   y = "Number of Items Ordered",
   caption = "Data from Instacart"
  ) 
```

**Problem 1, Part 4: Creating a table to show the 3 most popular items from each table**

```{r}
  instacart %>% 
  group_by(aisle, product_name) %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  summarize(n = n()) %>%
  top_n(n = 3) %>% 
  arrange(desc(n)) %>% 
  knitr::kable(caption = "3 Most Popular Items by Aisle")
```






   




