---
title: "p8105_hw3_tal2145"
author: Troy Layouni
date: 2019-10-08
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Loading packages**

```{r}
library(tidyverse)
library(ggridges)
```

## Problem 1

**Problem 1, Part 1: Loading Instacart data**

```{r}
library(p8105.datasets)
data("instacart")
```


**Problem 1, Part 2: Number of aisles and the aisle with the most items ordered**

```{r}
instacart_aisles = 
  instacart %>% 
  group_by(aisle) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n)) 

instacart_aisles
```

There are 134 aisles in the instacart dataset and the aisle that the most items are ordered from is the fresh vegetables. 

**Problem 1, Part 3: Plotting the number of items ordered in each aisle, limited to aisles with more than 10,000 items ordered**

```{r}
instacart_aisles %>% 
  filter(n > 10000) %>% 
  mutate(aisle = reorder(aisle, n)) %>% 
  
ggplot(aes(x = aisle, y = n)) +
  geom_col() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(
   title = "Number of Items Ordered by Aisle",
   x = "Aisle",
   y = "Number of Items Ordered",
   caption = "Data from Instacart"
  ) 
```

**Problem 1, Part 4: Creating a table to show the 3 most popular items from each table**

```{r}
  instacart %>% 
  group_by(product_name, aisle) %>% 
  summarize(n = n()) %>%
  #arrange(desc(n)) %>% 
  #top_n(n = 3) %>% 
  group_by(aisle) %>% 
  filter(
    aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits"), min_rank(desc(n)) < 4) %>% 
  arrange(n, aisle) %>% 
  knitr::kable(caption = "3 Most Popular Items by Aisle")
```

**Problem 1, Part 5: Creating a table for the mean hour of the day that both Pink Lady Apples and Coffee Ice Cream are ordered each day of the week ** 

```{r}
instacart %>% 
    mutate(order_dow = recode(order_dow, "0" = "Sunday", "1" = "Monday", "2" = "Tuesday", "3" = "Wednesday", "4" = "Thursday", "5" = "Friday", "6" = "Saturday")) %>% 
  group_by(product_name, order_dow) %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  mutate(mean_hour = mean(order_hour_of_day)) %>% 
  select(product_name, order_dow, mean_hour) %>% 
  distinct() %>% 
  pivot_wider(
    names_from = "order_dow",
    values_from = c("mean_hour")
  ) %>% 
  select(product_name, Sunday, Monday, Tuesday, Wednesday, Thursday, Friday, Saturday) %>% 
  knitr::kable(caption = "Instacart: Mean Hour Products are Bought by Day of Week")
```

## Problem 2

**Problem 2, Part 1: Loading data and data cleaning**

* Loading BRFSS from the `p8105.datasets` package
* Cleaning by: 
  * Formating to appropriate variable names
  * Focusing on the "Overall Health" topic
  * Including only responses from "Excellent" to "Poor"
  * Organizing responses as a factor taking levels ordered from "Poor" to "Excellent"
  
```{r}
library(p8105.datasets)
data("brfss_smart2010")

brfss_smart2010 =
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") 

brfss_smart2010 %>%  
  group_by(response) %>% 
  summarize(
    n_response = n()
  )
  
```

library(p8105.datasets)
data("brfss_smart2010")


  brfss_smart2010 = 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  group_by(response) %>% 
  summarize(
    n_question = n(response)
  )
  # mutate(response = as.factor(response)) %>%
  as.factor(response, levels =  c(""))


   




